---
title: "Bird Calls of Chernobyl"
author: "Sara Tamayo-Correa and Stephanie Tsai"
date: "Winter 2019"
output: 
  ioslides_presentation:
  widescreen: TRUE
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("jpeg")
```

## Packages Used
```{r}
library(tidyverse)
library(skimr)
library(paletteer)
```
## Loading Data
Database we are using is from "Kendrick, P.; Barçante, L.; Beresford, N.A.; Gashchak, S.; Wood, M.D. (2018). Bird Vocalisation Activity (BiVA) database: annotated soundscapes from the Chernobyl Exclusion Zone. NERC Environmental Information Data Centre. https://doi.org/10.5285/be5639e9-75e9-4aa3-afdd-65ba80352591"
Contains data supplied by Natural Environment Research Council from Open Government License

Clarification: "Buzz" represents insects

```{r}
call_data <- read_csv("C:/Users/Stephanie T/Desktop/FRS_417/stephanie_scripts_tests/complete_BirdCall_data_table.csv")
```
## Clean up data
Use of skimmr::skim to look if data may contain NAs or nonsense values

```{r}
skim(call_data)
```


## Renaming 
```{r}
call_data <- call_data %>% 
  rename(info = 'Additional information',
         sound_category = 'Sound category',
         begin_time_s = 'Begin Time (s)',
         end_time_s = 'End Time (s)',
         high_freq_hz = 'High Freq (Hz)',
         low_freq_hz = 'Low Freq (Hz)' ) 
head(call_data)

```

## Renaming the stored variable
```{r}
bird_calls <- call_data
```

Refining the columns we want to analyze
```{r}
bird_calls <- bird_calls %>% 
  select(begin_time_s, end_time_s, low_freq_hz, high_freq_hz, sound_category) 

head(bird_calls)
  
```

## Clean Data | All Animals included

```{r}
bird_call_clean <- bird_calls %>% 
  gather(low_freq_hz, high_freq_hz, key = "type", value = "frequency") %>% 
  mutate(duration = end_time_s - begin_time_s) %>% 
  mutate(animal_type = recode(sound_category, Buzz = "Insect")) #renames the default variable so the data is more clear
  

head(bird_call_clean)
```



#
What is the frequency range of each animal type?
```{r}
bird_call_clean %>% 
  ggplot(aes(x=animal_type, y=frequency, fill=animal_type))+
  geom_boxplot(position="dodge")+
  scale_fill_brewer(palette = "OrRd")+
  labs(title = "Frequency Range per Animal", x = "Animal Type", y = "Frequency (Hz)") +
  theme_light() +
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```

***
- Buzz (aka insects) had the lowest median of frequencies and the greatest outliers. 
- Birds have a median frequency of less than 5000 Hz and most of the frequencies are around that point. 
- We can not draw much of a conclusion from other animals because it is a mixture of all different unknown sounds.

## 
To further clarify what we are seeing, we divided these sounds by high and low frequency.
```{r}
bird_call_clean %>% 
  group_by(type) %>% 
  ggplot(aes(x = animal_type, y = frequency, fill = type)) +
  geom_boxplot() +
  labs(title = "Frequency by Animal",
       x = "Animal Type",
       y = "Frequency (Hz)",
       fill = "Frequency Type") +
  theme_bw() +
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
  
```

***
- This graph is unexpected because we expected more of the buzz sounds to be on the low frequency end
- It is hard to tell why this is the case, so we wil count how many data points make up the low frequency for insects.

##
Counting data points for low freqeuncy insects
```{r}
bird_call_clean %>% 
  filter(sound_category == "Buzz" & type == "low_freq_hz") %>% 
  ggplot(aes(x = frequency)) +
  geom_histogram(fill = "blue") +
  labs(x = "Frequency (Hz)", 
       y = "Count",
       title = "Low Frequency Count of Insect Sounds")+
  theme_bw() +
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```

***
- This tells us there are a lot of points on 0 Hz
- Our initial boxplot is misleading because it looks like there are not many points on low frequencies. 

## 
So is this the problem we see in all other animal types?
```{r}
bird_call_clean %>% 
  ggplot(aes(x=frequency))+
  geom_histogram(fill="lightyellow", color="black")+
  labs(title="Spread of Frequencies",
       x = "Frequency (Hz)",
       y = "Count")+
  theme_linedraw()+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```


## 
So what do the labels low frequency and high frequency mean?
```{r}
bird_call_clean %>% 
  ggplot(aes(x=type, y=frequency, fill=type))+
  geom_boxplot(position="dodge")+
  theme(legend.position = "right")+
  scale_fill_brewer(palette = "Pastel2")+
  labs(title = "Comparison of defined frequencies",
       x = "Frequency Type",
       y = "Frequency (Hz)",
       fill = "Frequency Type")+
  theme_light() +
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```

***
- Significant overlap between what is considered low and high frequency. 
- Because of this issue, the labels are possibly arbitrary. 
- The high frequency range dips into the low frequency and vice versa. 
- We cannot rely on their labels, so we will only use the raw frequency observations.

##
Total number of organisms recorded in Chernobyl per animal type
```{r}
bird_call_clean %>% 
  ggplot(aes(x = animal_type))+
  geom_bar(aes(fill = animal_type)) +
  scale_fill_paletteer_d(ggsci, legacy_tron) +
  theme_classic() +
  labs(title = "Number of individuals recorded in Chernobyl",
       x = "Animal",
       y = "Individuals",
       fill = "Animal Type") +
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```

***
- This tells us how many individuals were recorded. 
- There are many more birds than any other animal.



##
Is there a relationship between frequency range, recording duration, and animal type?
```{r}
bird_call_clean %>% 
  group_by(sound_category) %>% 
  ggplot(aes(x = duration, y = frequency)) +
  geom_point(alpha = 0.4, aes(color = frequency)) +
  scale_color_paletteer_c(harrypotter, newtscamander) +
  facet_wrap(~animal_type) +
  theme_classic() +
  labs(title = "Animal Call Frequency Duration",
       x = "Duration (seconds)",
       y = "Frequency (Hz)",
       color = "Frequency") +
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
 
  
```

***
- We see that birds have a frequency range between a little above 0-17500 Hz and have a call duration between about 0 and 100 seconds. 

- Insects have a slightly higher frequency range than birds, but they have shorter call durations. 

- Other animals have the greatest frequency range and can be divided into a low frequency or high frequency sound. They also have greater call durations than birds or insects.




















