---
title: "final picks"
author: "Stephanie Tsai and Sara TC"
date: "March 8, 2019"
output: 
  ioslides_presentation: 
    keep_md: yes
---

## Packages Used
```{r}
library(tidyverse)
library(skimr)
library(paletteer)
```
## Loading Data
Database we are using is from "Kendrick, P.; Barçante, L.; Beresford, N.A.; Gashchak, S.; Wood, M.D. (2018). Bird Vocalisation Activity (BiVA) database: annotated soundscapes from the Chernobyl Exclusion Zone. NERC Environmental Information Data Centre. https://doi.org/10.5285/be5639e9-75e9-4aa3-afdd-65ba80352591"
Contains data supplied by Natural Environment Research Council from Open Government License

-Clarification: "Buzz" represents insects

```{r}
call_data <- 
  readr::read_csv("~/Downloads/Bird_calls-master-2/complete_BirdCall_data_table.csv")
```
## Clean up data
-Use of skimmr::skim to look if data may contain NAs or nonsense values
```{r}
skim(call_data)
```
#Renaming 
```{r}
call_data <- call_data %>% 
  rename(info = 'Additional information',
         sound_category = 'Sound category',
         begin_time_s = 'Begin Time (s)',
         end_time_s = 'End Time (s)',
         high_freq_hz = 'High Freq (Hz)',
         low_freq_hz = 'Low Freq (Hz)' ) 


```

-Renaming the stored variable
```{r}
bird_calls <- call_data
```

#Refining the columns we want to analyze
```{r}
bird_calls <- bird_calls %>% 
  select(begin_time_s, end_time_s, low_freq_hz, high_freq_hz, sound_category) 

bird_calls
  
```

## Clean Data

-All Animals included
```{r}
bird_call_clean <- bird_calls %>% 
  gather(low_freq_hz, high_freq_hz, key = "type", value = "frequency") %>% 
  mutate(duration = end_time_s - begin_time_s) %>% 
  mutate(animal_type= recode(sound_category, Buzz= "Insect"))
  

bird_call_clean
```

 Questions
#1. What is the frequency range of each animal type?
```{r}
bird_call_clean %>% 
  ggplot(aes(x=sound_category, y=frequency, fill=animal_type))+
  geom_boxplot(position="dodge")+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))+
  scale_fill_brewer(palette = "OrRd")+
  labs(title = "Frequency")
```

-From this graph, we saw that Buzz (aka insects) had the lowest median of frequencies and the greatest outliers. Birds have a median frequency of less than 5000 Hz and most of the frequencies are around that point. We can not draw much of a conclusion from other animals because it is a mixture of all different unknown sounds.

#2. To further clarify what we are seeing, we divided these sounds by high and low frequency.
```{r}
bird_call_clean %>% 
  group_by(type) %>% 
  ggplot(aes(x = animal_type, y = frequency, fill = type)) +
  geom_boxplot() +
  labs(title = "Frequency by Animal",
       x = "Animal Type",
       y = "Frequency (Hz)",
       fill = "Frequency Type") +
  theme_bw()+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
  
```
-This graph is unexpected because we expected more of the buzz sounds to be on the low frequency end, but the data says that most are categorized on the high frequency end. It is hard to tell why this is the case, so we wil count how many data points make up the low frequency for insects.

#3. Counting data points for low freqeuncy insects
```{r}
bird_call_clean %>% 
  filter(animal_type == "Insect" & type == "low_freq_hz") %>% 
  ggplot(aes(x = frequency)) +
  geom_histogram(fill = "blue")+
  labs(title= "Frequency Count of Insect Sounds", x= "Frequency (Hz)", y= "Count")+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```
-This tells us there are a lot of points on 0 Hz, so our initial boxplot is misleading because it looks like there are not many points on low frequencies. 

#So is this the problem we see in all other animal types?
```{r}
bird_call_clean %>% 
  ggplot(aes(x=frequency))+
  geom_histogram(fill="lightyellow", color="black")+
  labs(title="The Spread of Frequency", x= "Frequency (Hz)", y="Count")+
  theme_linedraw()+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```


#4. So what do the labels low frequency and high frequency mean?
```{r}
bird_call_clean %>% 
  ggplot(aes(x=type, y=frequency, fill=type))+
  geom_boxplot(position="dodge")+
  theme(legend.position = "right")+
  scale_fill_brewer(palette = "Pastel2")+
  labs(title = "Comparison Defined Frequencies", x= "Frequency Type", y="Frequency (Hz)")+
  theme_light()+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
  
```
-Now we see there is significant overlap between what is considered low and high frequency. Because of this issue, the labels are possibly arbitrary. The high frequency range dips into the low frequency and vice versa. We cannot rely on their labels, so we will only use the raw frequency observations.

#5. Total number of organisms recorded in Chernobyl per animal type
```{r}
bird_call_clean %>% 
  ggplot(aes(x = animal_type))+
  geom_bar(aes(fill = sound_category)) +
  scale_fill_paletteer_d(ggsci, legacy_tron) +
  theme_classic() +
  labs(title = "Number of individuals recorded in Chernobyl",
       x = "Animal",
       y = "Individuals",
       fill = "Animal Type")+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
```
-This tells us how many individuals were recorded. There are many more birds than any other animal.



#6. Is there a relationship between frequency range, recording duration, and animal type?
```{r}
bird_call_clean %>% 
  group_by(animal_type) %>% 
  ggplot(aes(x = duration, y = frequency)) +
  geom_point(alpha = 0.4, aes(color = sound_category)) +
  scale_color_paletteer_d(ggsci, palette = category10_d3, direction = 1) +
  facet_wrap(~sound_category) +
  theme_light() +
  labs(title = "Animal Call Frequency Duration",
       x = "Duration (seconds)",
       y = "Frequency (Hz)",
       color = "Animal Type")+
  theme(plot.title = element_text(size=rel(1.5), hjust=0.5))
 
  
```

-We see that birds have a frequency range between a little above 0-17500 Hz and have a call duration between about 0-100 seconds.
[add interpretation here]











