---
title: "Animal Calls from Chernobyl"
author: "Stephanie Tsai"
date: "February 25, 2019"
output: 
  html_document: 
    keep_md: yes
---

## Packages Used
```{r}
library(tidyverse)
library(skimr)
library(paletteer)
```
## Loading Data
Database we are using is from "Kendrick, P.; Barçante, L.; Beresford, N.A.; Gashchak, S.; Wood, M.D. (2018). Bird Vocalisation Activity (BiVA) database: annotated soundscapes from the Chernobyl Exclusion Zone. NERC Environmental Information Data Centre. https://doi.org/10.5285/be5639e9-75e9-4aa3-afdd-65ba80352591"
Contains data supplied by Natural Environment Research Council from Open Government License

Clarification: "Buzz" represents insects

```{r}
call_data <- read_csv("C:/Users/Stephanie T/Desktop/FRS_417/stephanie_scripts_tests/complete_BirdCall_data_table.csv")
```
## Clean up data
Use of skimmr::skim to look if data may contain NAs or nonsense values
```{r}
skim(call_data)
```

Renaming 
```{r}
call_data <- call_data %>% 
  rename(info = 'Additional information',
         sound_category = 'Sound category',
         begin_time_s = 'Begin Time (s)',
         end_time_s = 'End Time (s)',
         high_freq_hz = 'High Freq (Hz)',
         low_freq_hz = 'Low Freq (Hz)' ) 


```

Renaming the stored variable
```{r}
bird_calls <- call_data
```

Refining the columns we want to analyze
```{r}
bird_calls <- bird_calls %>% 
  select(begin_time_s, end_time_s, low_freq_hz, high_freq_hz, sound_category) 

bird_calls
  
```


## Clean Data

All Animals included
```{r}
bird_call_clean <- bird_calls %>% 
  gather(low_freq_hz, high_freq_hz, key = "type", value = "frequency") %>% 
  mutate(duration = end_time_s - begin_time_s)
  

bird_call_clean
```
Only Birds
```{r}
bird_only <- bird_call_clean %>% 
  filter(sound_category == "Bird") 

bird_only
```


## Questions
1. Which category has the greatest frequency variation?
```{r}
bird_call_clean %>% 
  group_by(type) %>% 
  ggplot(aes(x = sound_category, y = frequency, fill = type)) +
  geom_boxplot() +
  labs(title = "Frequency by Animal",
       x = "Animal Type",
       y = "Frequency (Hz)",
       fill = "Frequency Type") +
  theme_bw()
  
```



2. Is there a relationship between frequency range and recording duration?
```{r}
bird_call_clean %>% 
  group_by(sound_category) %>% 
  ggplot(aes(x = duration, y = frequency)) +
  geom_point(alpha = 0.4, aes(color = sound_category)) +
  scale_color_paletteer_d(ggsci, palette = category10_d3, direction = 1) +
  facet_wrap(~sound_category) +
  theme_light() +
  labs(title = "Animal Call Frequency Duration",
       x = "Duration (seconds)",
       y = "Frequency (Hz)",
       color = "Animal Type")  #double check units
 
  
```

3. Trend across time
```{r}
bird_call_clean %>% 
  ggplot(aes(x = begin_time_s, y = frequency)) +
  geom_point(alpha = 0.4, aes(color = type)) +
  facet_wrap(~sound_category) +
  labs(title = "Frequency captured over time",
       x = "Time",
       y = "Frequency",
       color = "Frequency Type") +
  theme_light()
```


4. Total number of organisms recorded in Chernobyl
```{r}
bird_call_clean %>% 
  ggplot(aes(x = sound_category))+
  geom_bar(aes(fill = sound_category)) +
  scale_fill_paletteer_d(ggsci, legacy_tron) +
  theme_classic() +
  labs(title = "Number of individuals recorded in Chernobyl",
       x = "Animal",
       y = "Individuals",
       fill = "Animal Type")
```

5. Bird Call Frequency with respect to time and duration of call
```{r}
bird_only %>% 
  filter(type == "low_freq_hz") %>% 
  ggplot(aes(x = begin_time_s, y = frequency, color = duration)) +
  geom_point(alpha = 0.6) +
  scale_color_paletteer_c("harrypotter", "ronweasley", direction = -1) +
  theme_bw()+
  labs(title = "Low Frequency Duration over time",
       x = "Time (s)",
       y = "Frequency",
       color = "Duration of Call (s)")
```




```{r}
bird_only %>% 
  filter(type == "high_freq_hz") %>% 
  ggplot(aes(x = begin_time_s, y = frequency, color = duration)) +
  geom_point(alpha = 0.6) +
  scale_color_paletteer_c("viridis", "plasma", direction = -1) +
  theme_bw()+
  labs(title = "High Frequency Duration over time",
       x = "Time (s)",
       y = "Frequency",
       color = "Duration of Call (s)")
```


```{r}
bird_only %>% 
  
  ggplot(aes(x = begin_time_s, y = duration, color = frequency)) +
  geom_point(alpha = 0.3) +
  scale_color_paletteer_c("pals", "kovesi.rainbow_bgyr_35_85_c72", direction = 1, ) +
  theme_bw()+
  labs(title = "Frequency over time",
       x = "Time (s)",
       y = "Duration",
       color = "Frequency")
```



